---
title: "DFAmethods2"
author: "Ikaro Daniel de Carvalho Barreto, Luiz Henrique Dore e Antonio Samuel Alves da Silva"
date: "03/09/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DFAmethods2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
**DFAMethods2** is a package for Detrended Fluctuation like-Analysis. In this package we developed methods such as:

**Detrended Fluctuation Analysis**  
**Detrended Cros-Correlation Analysis**  
**Detrended Cros-Correlation Coefficient**  
**Detrended Partial Cros-Correlation Coefficient**  
**Detrended Multiple Linear Regression**  
**Beta DFA estimates**  
**Standardized Beta DFA estimates**  
**f² scale-wise effect size**  
**Shen-Podobnik Test**  
**Kristoufek Test**  
**Intersection-Union Test**  

```{r setup}
library(DFAmethods2)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Usage

This section provides examples of how to use the main functions of the `DFAmethods2` package. We will use the `EuStockMarkets` dataset, which is included in R.

```{r data}
# Load the dataset
data(EuStockMarkets)

# Convert to a data.frame
df <- as.data.frame(EuStockMarkets)

# Use the first 1000 observations for this example
df_subset <- df[1:1000, ]
```

## 1. Detrended Fluctuation Analysis (DFA)

The `dfa()` function calculates the Detrended Fluctuation Analysis for a single time series.

```{r dfa_example}
# Perform DFA on the DAX index
dfa_result <- dfa(df_subset$DAX)

# Print the first few rows of the result
head(dfa_result)
```

## 2. Detrended Cross-Correlation Analysis (DCCA)

The `rhodcca()` function calculates the Detrended Cross-Correlation Coefficient for two or more time series.

```{r rhodcca_example}
# Perform DCCA on the DAX and SMI indices
rhodcca_result <- rhodcca(df_subset[, c("DAX", "SMI")])

# Print the first few rows of the result
head(rhodcca_result)
```

## 3. Detrended Partial Cross-Correlation Analysis (DPCCA)

The `rhodpcca()` function calculates the Detrended Partial Cross-Correlation Coefficient for three or more time series.

```{r rhodpcca_example}
# Perform DPCCA on DAX, SMI, and CAC indices
rhodpcca_result <- rhodpcca(df_subset[, c("DAX", "SMI", "CAC")])

# Print the first few rows of the result
head(rhodpcca_result)
```

## 4. Detrended Multiple-Correlation Coefficient (DMC²)

The `dmc2()` function calculates the Detrended Multiple-Correlation Coefficient.

```{r dmc2_example}
# Calculate DMC² for DAX, SMI, and CAC
dmc2_result <- dmc2(df_subset[, c("DAX", "SMI", "CAC")])

# Print the first few rows of the result
head(dmc2_result)
```

## 5. Multiple Fractal Regression

The `fracreg()` function calculates the Multiple Fractal Regression.

```{r fracreg_example}
# Perform fractal regression
fracreg_result <- fracreg(df_subset, dpo = 1, int = TRUE)

# The result is a list of objects. For example, you can access the Beta DFA estimates
head(fracreg_result$BDFA)
```

