---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DFAmethods2

<!-- badges: start -->
<!-- badges: end -->

**DFAmethods2** is a toolbox of Detrended Fluctuation Analysis (DFA) methods for
measuring long-range correlations, cross-correlations and regression in
**nonstationary** time series, where classical autocorrelation and ordinary
regression are biased by trends. It provides:

- **Detrended Fluctuation Analysis** — `dfa()`
- **Detrended Cross-Correlation Analysis** and the **rho-DCCA** coefficient — `rhodcca()`
- **Detrended Partial Cross-Correlation** coefficient — `rhodpcca()`
- **Detrended Multiple Cross-Correlation** coefficient (DMC) — `dmc2()`
- **Detrended fractal regression**: scale-wise and standardized coefficients — `betadfa()`, `sbdfa()`, `fracreg()`
- **Scale-wise** $f^2$ **effect sizes** — `effsizeDFA()`
- **Significance tests**: Podobnik-Shen, Kristoufek and intersection-union — `fracreg.PStest()`, `fracreg.Ktest()`, `fracreg.IUTest()`

The scale-dependent standardized coefficients, the scale-wise effect size and
the intersection-union test follow Barreto et al. (2021)
[doi:10.1016/j.physa.2021.126259](https://doi.org/10.1016/j.physa.2021.126259).

## Installation

``` r
# install.packages("devtools")
devtools::install_github("Ikarobarreto/DFAmethods2")
```

## Example

A response `y` driven by a covariate `x2`, both random-walk-like:

```{r example}
library(DFAmethods2)

set.seed(1)
x2  <- cumsum(rnorm(800))
y   <- 0.7 * x2 + cumsum(rnorm(800))
dat <- data.frame(y = y, x2 = x2)

# DFA fluctuation function of y
head(dfa(y, np = 30))

# scale-wise rho-DCCA cross-correlation coefficient
head(rhodcca(dat, np = 30))
```

See `vignette("DFAmethods2")` for the full theory and worked examples.
