---
title: "DFAmethods2"
author: "Ikaro Daniel de Carvalho Barreto, Luiz Henrique Dore e Antonio Samuel Alves da Silva"
date: "03/09/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DFAmethods2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
**DFAMethods2** is a package for Detrended Fluctuation like-Analysis. In this package we developed methods such as:

**Detrended Fluctuation Analysis**  
**Detrended Cros-Correlation Analysis**  
**Detrended Cros-Correlation Coefficient**  
**Detrended Partial Cros-Correlation Coefficient**  
**Detrended Multiple Linear Regression**  
**Beta DFA estimates**  
**Standardized Beta DFA estimates**  
**f² scale-wise effect size**  
**Shen-Podobnik Test**  
**Kristoufek Test**  
**Intersection-Union Test**  

```{r setup}
library(DFAmethods2)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# 1. Detrended Fluctuation Analysis

The Detrended Fluctuation Analysis (DFA) is a method used to quantify long-range correlations in time series.

```{r}
# Generate a sample time series
data <- rnorm(1000)

# Calculate DFA
dfa_result <- dfa(data)

# Print the first few rows of the result
head(dfa_result)
```

# 2. Detrended Cros-Correlation Analysis

The Detrended Cross-Correlation Analysis (DCCA) is a method to quantify long-range cross-correlations between two non-stationary time series.

```{r}
# Generate a sample matrix of two time series
data <- matrix(rnorm(2000), ncol = 2)

# Calculate DCCA using the fracreg function
dcca_result <- fracreg(data, dpo = 1, int = TRUE)

# The DCCA results are in the 'F' component of the list
head(dcca_result$F)
```

# 3. Detrended Cros-Correlation Coefficient

The Detrended Cross-Correlation Coefficient (rho-DCCA) is a measure of the strength of cross-correlation.

```{r}
# Generate a sample matrix of two time series
data <- matrix(rnorm(2000), ncol = 2)

# Calculate rho-DCCA
rhodcca_result <- rhodcca(data)

# Print the first few rows of the result
head(rhodcca_result)
```

# 4. Detrended Partial Cros-Correlation Coefficient

The Detrended Partial Cross-Correlation Coefficient (rho-DPCCA) is a measure of the strength of cross-correlation between two time series after removing the effect of other time series.

```{r}
# Generate a sample matrix of three time series
data <- matrix(rnorm(3000), ncol = 3)

# Calculate rho-DPCCA
rhodpcca_result <- rhodpcca(data)

# Print the first few rows of the result
head(rhodpcca_result)
```

# 5. Detrended Multiple-Correlation Coefficient

The `dmc2` function calculates the Detrended Multiple-Correlation Coefficient.

```{r}
# Generate a sample matrix of three time series
data <- matrix(rnorm(3000), ncol = 3)

# Calculate Detrended Multiple-Correlation Coefficient
dmc2_result <- dmc2(data)

# Print the first few rows of the result
head(dmc2_result)
```

# 6. Detrended Multiple Linear Regression

The `fracreg` function performs a Detrended Multiple Linear Regression.

```{r}
# Generate a sample matrix of three time series
data <- matrix(rnorm(3000), ncol = 3)

# Calculate Detrended Multiple Linear Regression
fracreg_result <- fracreg(data, dpo = 1, int = TRUE)

# The results are a list of various components. For example, the Beta DFA estimates:
head(fracreg_result$BDFA)
```

# 7. Beta DFA estimates

The `betadfa` function calculates the Beta DFA estimates.

```{r}
# Generate a sample matrix of three time series
data <- matrix(rnorm(3000), ncol = 3)

# Calculate Beta DFA estimates
betadfa_result <- betadfa(data)

# Print the first few rows of the result
head(betadfa_result)
```

# 8. Standardized Beta DFA estimates

The `sbdfa` function calculates the Standardized Beta DFA estimates.

```{r}
# Generate a sample matrix of three time series
data <- matrix(rnorm(3000), ncol = 3)

# Calculate Standardized Beta DFA estimates
sbdfa_result <- sbdfa(data)

# Print the first few rows of the result
head(sbdfa_result)
```

# 9. f² scale-wise effect size

The `effsizeDFA` function calculates the f² scale-wise effect size.

```{r}
# Generate a sample matrix of three time series
data <- matrix(rnorm(3000), ncol = 3)

# Calculate f² scale-wise effect size
effsizeDFA_result <- effsizeDFA(data)

# Print the first few rows of the result
head(effsizeDFA_result)
```

# 10. Shen-Podobnik Test

The `fracreg.PStest` function performs the Podobnik-Shen Test.

```{r}
# Generate a sample matrix of two time series
data <- matrix(rnorm(2000), ncol = 2)

# Perform the Podobnik-Shen Test
PStest_result <- fracreg.PStest(data)

# Print the first few rows of the result
head(PStest_result)
```

# 11. Kristoufek Test

The `fracreg.Ktest` function performs the Kristoufek Test.

```{r}
# Generate a sample matrix of two time series
data <- matrix(rnorm(2000), ncol = 2)

# Perform the Kristoufek Test
Ktest_result <- fracreg.Ktest(data)

# Print the first few rows of the result
head(Ktest_result)
```

# 12. Intersection-Union Test

The `fracreg.IUTest` function performs the Intersection-Union Test.

```{r}
# Generate a sample matrix of two time series
data <- matrix(rnorm(2000), ncol = 2)

# Perform the Intersection-Union Test
IUTest_result <- fracreg.IUTest(data)

# The result is a list containing the results of the PStest and Ktest
names(IUTest_result)
head(IUTest_result$PStest)
head(IUTest_result$Ktest)
```

# 13. Plotting Functions

## 13.1 plotdfa

The `plotdfa` function plots the results of a DFA analysis.

```{r}
# Generate a sample time series
data <- rnorm(1000)

# Calculate DFA
dfa_result <- dfa(data)

# Plot the DFA result
plotdfa(dfa_result, main = "DFA Plot")
```

## 13.2 plotdcca

The `plotdcca` function plots the results of a DCCA analysis.

```{r}
# Generate a sample matrix of two time series
data <- matrix(rnorm(2000), ncol = 2)

# Calculate DCCA using the fracreg function
dcca_result <- fracreg(data, dpo = 1, int = TRUE)

# Plot the DCCA result
plotdcca(dcca_result, main = "DCCA Plot")
```

## 13.3 plotrdcca

The `plotrdcca` function plots the results of a rho-DCCA analysis.

```{r}
# Generate a sample matrix of two time series
data <- matrix(rnorm(2000), ncol = 2)

# Calculate rho-DCCA
rdcca_result <- rhodcca(data)

# Plot the rho-DCCA result
plotrdcca(rdcca_result, var = 1, main = "rho-DCCA Plot")
```
