---
title: "DFAmethods2"
author: "Ikaro Daniel de Carvalho Barreto, Luiz Henrique Dore e Antonio Samuel Alves da Silva"
date: "03/09/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DFAmethods2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
**DFAMethods2** is a package for Detrended Fluctuation like-Analysis. In this package we developed methods such as:

**Detrended Fluctuation Analysis**  
**Detrended Cros-Correlation Analysis**  
**Detrended Cros-Correlation Coefficient**  
**Detrended Partial Cros-Correlation Coefficient**  
**Detrended Multiple Linear Regression**  
**Beta DFA estimates**  
**Standardized Beta DFA estimates**  
**f² scale-wise effect size**  
**Shen-Podobnik Test**  
**Kristoufek Test**  
**Intersection-Union Test**  

```{r setup}
library(DFAmethods2)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# 1. Detrended Fluctuation Analysis

# 2. Detrended Cros-Correlation Analysis

# 3. Detrended Cros-Correlation Coefficient

# 4. Detrended Partial Cros-Correlation Coefficient

# 5. Detrended Multiple Linear Regression

# 6. Beta DFA estimates

# 7. Standardized Beta DFA estimates

# 8. f² scale-wise effect size

# 9. Shen-Podobnik Test

# 10. Kristoufek Test

# 11. Intersection-Union Test

