---
title: "Detrended fluctuation analysis and related methods"
author: "Ikaro Daniel de Carvalho Barreto, Luiz Henrique Dore"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Detrended fluctuation analysis and related methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 6,
  fig.height = 4,
  fig.align = "center"
)
set.seed(42)
```

## Overview

Detrended Fluctuation Analysis (DFA) and its relatives are a family of methods
for measuring long-range correlations and cross-correlations in **nonstationary**
time series, where classical autocorrelation and ordinary regression are biased
by trends. `DFATools` collects the most useful members of this family under a
common, consistent interface:

| Concern | Function(s) |
|---|---|
| Scaling of a single series | `dfa()`, `plotdfa()` |
| Cross-correlation of two series | `rhodcca()`, `plotrdcca()` |
| Partial cross-correlation | `rhodpcca()` |
| Multiple cross-correlation | `dmc2()` |
| Regression at each scale | `betadfa()`, `sbdfa()`, `fracreg()` |
| Scale-wise effect sizes | `effsizeDFA()` |
| Coefficient inference, diagnostics & robust SEs | `fracreg()`, `fracreg.diag()`, `fracreg.WB()` |
| Significance tests | `fracreg.PStest()`, `fracreg.Ktest()`, `fracreg.IUTest()` |

Several of the regression-oriented tools — the scale-dependent **standardized**
coefficients, the scale-wise $f^2$ **effect size** and the **intersection-union
test** — were introduced by @barreto2021, whose framework this package
implements.

```{r setup}
library(DFATools)
```

### Data convention and common arguments

Except for `dfa()` (which takes a single numeric vector), the functions expect a
**matrix or data frame whose first column is the response** $Y$ and whose
remaining columns are the covariates $X_2, \dots, X_p$. They share four
arguments:

* `dpo` — order of the polynomial used for detrending (1 = linear);
* `int` — if `TRUE`, the series are integrated (cumulatively summed) first;
* `np` — number of scales (box sizes) at which the statistics are evaluated;
* `overlap` — if `TRUE`, sliding (overlapping) boxes are used.

Throughout the vignette we use a simulated system in which a response `y`
depends on two random-walk drivers, `x2` (positive) and `x3` (negative):

```{r data}
n  <- 1200
x2 <- cumsum(rnorm(n))
x3 <- cumsum(rnorm(n))
y  <- 0.7 * x2 - 0.5 * x3 + cumsum(rnorm(n))
dat <- data.frame(y = y, x2 = x2, x3 = x3)
```

## Detrended fluctuation analysis

DFA [@peng1994] quantifies the scaling of fluctuations in a series. The series
is first integrated into the profile $X(t) = \sum_{i=1}^{t} (x_i - \bar{x})$,
which is split into boxes of size $s$; within each box a polynomial trend is
removed and the variance of the residuals is taken. Averaging over boxes gives
the **fluctuation function** $F(s)$, which for a self-similar series follows a
power law
$$
F(s) \sim s^{\alpha},
$$
where the exponent $\alpha$ generalises the Hurst exponent: $\alpha = 0.5$ marks
an uncorrelated series, $\alpha > 0.5$ persistence (long-range positive
correlations) and $\alpha < 0.5$ anti-persistence.

```{r dfa}
fy <- dfa(dat$y, np = 40)
head(data.frame(s = fy$s, F = fy$F, F2 = fy$F2))

# the DFA exponent is the slope of log F(s) against log s,
# also returned ready-to-use as fy$alpha:
round(c(slope = coef(lm(log10(fy$F) ~ log10(fy$s)))[[2]],
        alpha = fy$alpha), 3)
```

`plotdfa()` draws the log–log fluctuation plot with the fitted exponent:

```{r dfa-plot, fig.alt = "DFA log-log fluctuation plot"}
plotdfa(fy, main = "DFA of y")
```

## Cross-correlation: the rho-DCCA coefficient

Detrended Cross-Correlation Analysis (DCCA) [@podobnik2008] replaces the box
variance by the **detrended covariance** $F_{XY}^2(s)$ between two integrated
profiles. On its own $F_{XY}^2(s)$ is unbounded, so @zebende2011 normalised it
into the **DCCA cross-correlation coefficient**
$$
\rho_{\mathrm{DCCA}}(s) = \frac{F_{XY}^2(s)}{F_X(s)\,F_Y(s)} \in [-1, 1],
$$
a scale-by-scale analogue of Pearson's correlation that is robust to trends.
`rhodcca()` returns one column per pair of series:

```{r rhodcca}
rc <- rhodcca(dat, np = 40)
head(rc)
```

```{r rhodcca-plot, fig.alt = "rho-DCCA coefficient against scale"}
plotrdcca(rc, var = "12")   # pair y (1) vs x2 (2)
```

The coefficient is strongly positive for the `y`–`x2` pair, as built into the
simulation.

## Partial cross-correlation: rho-DPCCA

When several series interact, a direct $\rho_{\mathrm{DCCA}}$ may be inflated by
shared drivers. Detrended Partial-Cross-Correlation Analysis [@yuan2015] removes
the linear influence of the remaining series by inverting the
$\rho_{\mathrm{DCCA}}$ matrix at each scale, giving the **partial** coefficient
$\rho_{\mathrm{DPCCA}}(s)$:

```{r rhodpcca}
rp <- rhodpcca(dat, np = 40)
head(rp)
```

## Multiple cross-correlation: DMC

The Detrended Multiple Cross-Correlation Coefficient [@zebende2018] generalises
the multiple-correlation coefficient ($R$) to scales. Writing
$\boldsymbol{\rho}(s)$ for the matrix of $\rho_{\mathrm{DCCA}}(s)$ among the
covariates and $\boldsymbol{\rho}_{Y\cdot}(s)$ for their coefficients with the
response,
$$
\mathrm{DMC}(s) = \boldsymbol{\rho}_{Y\cdot}(s)^{\top}\,
                  \boldsymbol{\rho}(s)^{-1}\,
                  \boldsymbol{\rho}_{Y\cdot}(s) \in [0, 1],
$$
measuring how much of the response is jointly explained by the covariates at
each scale (a recent statistical test is given by @wang2021):

```{r dmc}
dm <- dmc2(dat, np = 40)
plot(dm[[1]], dm[[2]], type = "b", pch = 19, ylim = c(0, 1),
     xlab = "scale s", ylab = "DMC(s)", main = "Detrended multiple correlation")
```

## Detrended fractal regression

@kristoufek2015 turned the detrended (co)variances into a full **regression
framework**. Collecting the detrended covariance matrix $\mathbf{F}(s)$, the
scale-wise regression coefficients are
$$
\boldsymbol{\beta}(s) = \mathbf{F}_{XX}(s)^{-1}\,\mathbf{F}_{XY}(s),
$$
an unstandardised effect of each covariate on the response **at scale $s$** that,
unlike OLS, can vary across scales. The multivariate case ($p \ge 2$ covariates,
detrended multiple linear regression) is due to @shen2015. `betadfa()` returns
these coefficients and `sbdfa()` their **standardised** counterparts (computed
from the $\rho_{\mathrm{DCCA}}$ matrices), which @barreto2021 proposed as a
scale-wise measure of relative importance:

```{r betadfa}
b  <- betadfa(dat, np = 40)
head(b)

matplot(b$s, b[, c("x2", "x3")], type = "l", lty = 1, lwd = 2,
        xlab = "scale s", ylab = expression(beta(s)),
        main = "Scale-wise regression coefficients")
abline(h = c(0.7, -0.5), lty = 3, col = "grey50")
legend("right", c("x2", "x3"), col = 1:2, lty = 1, lwd = 2, bty = "n")
```

The single-path estimates sit near the true coefficients $0.7$ and $-0.5$, but a
single realisation cannot establish unbiasedness --- averaging over scales does
not shrink the sampling error of one path. A small Monte Carlo experiment with
**stationary** predictors makes the recovery rigorous, reporting the mean and the
Monte Carlo standard error (MCSE) over replications:

```{r mc}
set.seed(1)
n <- 1500; nsim <- 100
b <- c(x2 = 0.7, x3 = -0.5)
est <- t(replicate(nsim, {
  x2 <- rnorm(n); x3 <- rnorm(n)
  d  <- data.frame(y = b["x2"] * x2 + b["x3"] * x3 + rnorm(n, sd = 0.5),
                   x2 = x2, x3 = x3)
  fit <- fracreg(d, dpo = 1, int = TRUE, np = 30, overlap = FALSE)
  ce  <- which(fit$s >= 20 & fit$s <= n / 10)        # central scales
  c(mean(fit$BDFA[1, 1, ce]), mean(fit$BDFA[2, 1, ce]))
}))
round(rbind(true = b,
            mean = colMeans(est),
            mcse = apply(est, 2, sd) / sqrt(nsim)), 3)
```

Both coefficients are recovered to within a couple of Monte Carlo standard errors
of their true values.

`fracreg()` is the all-in-one workhorse: a single call returns the scale vector,
detrended (partial) correlations, raw and standardised $\beta(s)$, the
coefficient of determination, the DMC, confidence limits and $p$-values:

```{r fracreg}
fr <- fracreg(dat, dpo = 1, int = TRUE, np = 40)
names(fr)
```

### Scale-wise effect sizes

`effsizeDFA()` reports the scale-wise effect size introduced by @barreto2021 —
Cohen's $f^2$ [@cohen1988] evaluated at each scale,
$f^2 = (R^2_{\text{full}} - R^2_{\text{reduced}}) / (1 - R^2_{\text{full}})$ —
turning the regression into a scale-resolved importance measure:

```{r effsize}
ef <- effsizeDFA(dat, np = 40)
head(ef)
```

## Inference: confidence intervals, diagnostics and robust standard errors

The point estimates above are only half the story. `fracreg()` also quantifies
the **uncertainty** of every scale-wise coefficient: the returned arrays carry
the variance (`VDFA`), the 95% confidence limits (`LCIB`, `UCIB`) and two-sided
`p.value`s, alongside collinearity diagnostics. A coefficient is indexed as
covariate $j$ at scale $k$ — e.g. `fr$BDFA[j, 1, k]` and `fr$UCIB[j, 1, k]`.

### The coefficient variance

By default (`variance = "inv_corrected"`) the variance is built from the **full
inverse** of the predictors' detrended covariance matrix, normalised by the box
count $T_s = \lfloor N/s\rfloor$ minus the number of predictors $k$, and scaled
by a memory-correction factor,
$$
\operatorname{var}\!\big(\hat\beta_j(s)\big)
   = \frac{F^2_{\varepsilon}(s)\,\big[\mathbf{F}_{XX}(s)^{-1}\big]_{jj}}{T_s - k}
   \cdot \frac{(2\widehat H + 1)^2}{21}.
$$
The inverse is the matrix that produces $\boldsymbol{\beta}(s)$ [@tilfani2022];
the factor depends on the DFA exponent $\widehat H$ of the regression error
(estimated from the OLS residual) and restores nominal coverage under polynomial
detrending. `variance = "inv"` drops the factor, and the legacy
`variance = "marginal"` form $F^2_{\varepsilon}(s)/F^2_{X_j}(s)$ ignores the
off-diagonal structure and **under-covers when the covariates are correlated**.

```{r ci, fig.alt = "x2 coefficient with 95 percent confidence band across scales"}
fr <- fracreg(dat, dpo = 1, int = TRUE, np = 40)   # variance = "inv_corrected" (default)
b  <- fr$BDFA[1, 1, ]                               # x2 coefficient
lo <- fr$LCIB[1, 1, ]; hi <- fr$UCIB[1, 1, ]
plot(fr$s, b, type = "n", ylim = range(lo, hi, na.rm = TRUE),
     xlab = "scale s", ylab = expression(beta[x2](s)),
     main = "x2 coefficient with 95% confidence band")
polygon(c(fr$s, rev(fr$s)), c(lo, rev(hi)), col = "grey85", border = NA)
lines(fr$s, b, lwd = 2)
abline(h = 0.7, lty = 3, col = "grey40")
```

### Collinearity diagnostics

Because the variance now depends on the conditioning of $\mathbf{F}_{XX}(s)$,
`fracreg()` reports scale-wise diagnostics: the variance inflation factors
`VIF`, the condition number `kappa` of the covariates' scale-wise correlation
matrix, and the adjusted coefficient of determination `R2adj` [@tilfani2022].
`fracreg.diag()` returns them as a tidy table:

```{r diag}
head(fracreg.diag(dat, dpo = 1, int = TRUE, np = 40))
```

Here the two drivers are independent, so the `VIF`s sit at 1 and `kappa` near 1;
a `VIF` rising well above 1 at some scales would flag scale-localised
collinearity that inflates the coefficient variance — a warning with no
analogue in ordinary regression.

### Robust standard errors: HC and the wild bootstrap

The analytic variance assumes homoscedastic, independent detrended boxes. When
that is in doubt, two robustifications are available, both built on the per-box
detrended moments (so the DFA itself is never recomputed):

* `variance = "hc"` — a heteroscedasticity-consistent (sandwich) variance,
  $\mathbf{F}_{XX}^{-1}\boldsymbol{\Omega}\,\mathbf{F}_{XX}^{-1}$ with
  $\boldsymbol{\Omega}$ the per-box outer product of the moment scores. The
  coefficients are unchanged — only their variance is. Use `overlap = FALSE`.
* `fracreg.WB()` — a **wild bootstrap** of the coefficients that resamples those
  scores and returns percentile confidence limits and $p$-values. The default
  `weights = "dependent"` is the dependent wild bootstrap of @shao2010 (robust
  to dependence across boxes); `"rademacher"` and `"mammen"` [@mammen1993] give
  the i.i.d. version, which reproduces the HC variance exactly.

```{r hc}
fr_hc <- fracreg(dat, dpo = 1, int = TRUE, np = 40, overlap = FALSE, variance = "hc")
# HC standard error of the x2 coefficient (first few scales)
round(sqrt(fr_hc$VDFA[1, 1, ])[1:6], 3)
```

```{r wb, fig.alt = "x2 coefficient with wild-bootstrap confidence band"}
wb <- fracreg.WB(dat, B = 199, weights = "dependent", np = 20)
plot(wb$s, wb$beta_x2, type = "n", ylim = range(wb$lower_x2, wb$upper_x2, na.rm = TRUE),
     xlab = "scale s", ylab = expression(beta[x2](s)),
     main = "x2 coefficient with wild-bootstrap 95% band")
polygon(c(wb$s, rev(wb$s)), c(wb$lower_x2, rev(wb$upper_x2)),
        col = "grey85", border = NA)
lines(wb$s, wb$beta_x2, lwd = 2)
abline(h = 0.7, lty = 3, col = "grey40")
```

Both the analytic and the bootstrap bands keep `x2` clearly above zero at every
scale, confirming a robust, scale-consistent positive effect.

## Significance tests

`DFATools` ships surrogate-based tests for the scale-wise coefficients, built
on phase-randomised surrogates of the series:

* `fracreg.PStest()` — Podobnik–Shen test of $H_0\!: \beta(s) = 0$ (no
  detrended effect) [@podobnik2011; @shen2015];
* `fracreg.Ktest()` — Kristoufek test of $H_0\!: \beta_{\mathrm{DFA}}(s) =
  \beta_{\mathrm{OLS}}$ (the scale effect equals the global OLS one)
  [@kristoufek2015];
* `fracreg.IUTest()` — the intersection–union test of @barreto2021 (following
  the intersection–union principle, e.g. @vandeun2009), which combines the two
  so as to reject only when a covariate has a direct, scale-dependent effect
  that OLS would miss.

Each returns the estimated $\beta(s)$ together with the lower/upper critical
bounds of the test, so a coefficient is significant at a scale when it falls
outside the band. We use a small number of surrogates here for speed; raise `B`
in practice.

```{r tests}
ps <- fracreg.PStest(dat, B = 30, np = 30)
head(ps)
```

```{r tests-plot, fig.alt = "Scale-wise beta with Podobnik-Shen critical band"}
plot(ps$s, ps$bet1, type = "l", lwd = 2,
     ylim = range(ps[, c("bet1", "slci1", "suci1")]),
     xlab = "scale s", ylab = expression(beta[x2](s)),
     main = "x2 coefficient vs Podobnik-Shen null band")
lines(ps$s, ps$slci1, lty = 2, col = "red")
lines(ps$s, ps$suci1, lty = 2, col = "red")
```

The `x2` coefficient lies well above the null band at all scales — a clear,
scale-consistent effect.

## Practical notes

* Choose `np` and the implied range of box sizes with care: very small boxes are
  noisy and very large boxes leave few windows. The defaults span box sizes from
  10 to one fifth of the series length.
* Use `int = TRUE` for noise-like (stationary) inputs and `int = FALSE` when the
  series are already integrated (random-walk-like).
* For confidence intervals, keep the default `variance = "inv_corrected"`; switch to the
  wild bootstrap (`fracreg.WB()`) when you suspect heteroscedasticity or
  dependence between boxes, and read the `VIF`/`kappa` diagnostics to spot scales
  where collinearity inflates the variance.
* `fracreg()` and `fracreg.WB()` accept `abs = TRUE` to build the
  cross-correlation step from the **absolute** detrended covariance, which is
  more robust to outliers.
* The surrogate tests are computationally heavy (they refit the model `B` times);
  start small and increase `B` once the workflow is settled.

## References
